---
title: "Introduction to fars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
### 2023-07-11
***
## Overview
The fars collection of functions are used to ingest and analyze raw data downloaded from the Fatality Analysis Reporting System hosted on the National Highway Traffic Safety Administration (NHTSA) website:

https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars

The fars collection includes the following functions:

* `fars_read()`
* `fars_read_years()`
* `fars_summarize_years()`
* `fars_map_state()`
* `make_filename()`

#### Dependencies
>The following packages are required to use the fars function:
`readr()`, `dplyer()`, `tidyr()`, `maps()` and `graphics()`.

***
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(fars)
```

## FARS Data
The fars functions read data from the *working directory* in the format of the
FARS: `r make_filename(2013)`. Each file represents a year of accident data in 
a compressed CSV. The fars helper function, `make_filename()`, takes a year as
an input parameter and generates the aforementioned file format. This output is
used by many of the other fars functions.

## `fars_read()`
This function checks if the specified file exists, reads the CSV file into a 
data frame, suppresses any messages, and converts the data frame into a tibble
before returning it as the output.
```{r fars_read}
fars_read("accident_2014.csv.bz2")
```

## `fars_read_years()`
This function reads and processes data files in the working directory for 
multiple years; it takes a vector of numeric years as input and returns a 
list of processed data frames.
```{r fars_read_years}
fars_read_years(c(2013,2014,2015))
```

## `fars_summarize_years()`
This function summarizes the data obtained from multiple years by counting the 
number of records for each month in each year. The function takes a vector of 
years as input and returns a data frame with summarized information.
```{r fars_summarize_years}
fars_summarize_years(2013:2015)
```

## `fars_map_state()`
This function plots the locations of accidents on a map for a specific state and year using data from the Fatality Analysis Reporting System (FARS). The function takes two parameters: `state.num`, which represents the state number and `year`, which represents the year of the data.
```{r fars_map_state, fig.dim=c(7,7)}
# Plot 2013 FARS data for the state of Texas
fars_map_state(48, 2013)
```

